### 2021-08-19 documentos/notas/notas_ja.tex
Reset error count between rows. herpderp
diagonalización de la matriz de Choi de qutrits con GellMann

\section{CP from eigenvalues} % {{{
% Intro {{{
Choi matrix's eigenvalues of a 1-qubit Pauli Channel are
\begin{align}
\lambda_{\mu}=1+\sum_{j=1}^3a_j^{\mu}\tau_j
\hspace{35pt}
\mu=0,1,2,3;
\hspace{35pt}
a_j^{\mu}=\left\{ \begin{array}{rl}
             1, & \mu=j \lor \mu=0\\
             -1, & \mu\neq j
             \end{array}
   \right..
\end{align}
Therefore, the inequalities needed to be sattisfied for a 1-qubit 
unital Pauli channel in order to be CP may be expressed in the 
following compact expression:
\begin{equation}
\sum_{j=0}^3a_j^{\mu}\tau_j\geq -1,
\hspace{35pt}
\mu=1,2,3;
\hspace{35pt}
a_j^{\mu}=\left\{ \begin{array}{rl}
             1, & \mu=j\\
             -1, & \mu\neq j
             \end{array}
   \right..
\end{equation}

@@

Therefore, the inequalities needed to be sattisfied for the CP are
\begin{equation}\label{eq:inequalities-PCE}
\sum_{\underset{(j_1,\ldots,j_n\neq0)}{j_1,\ldots,j_n=0}}^3
a_{j_1}^{\mu_1}\ldots a_{j_n}^{\mu_n}\tau_{j_1,\ldots,j_n}
\geq
-1,
\hspace{35pt}
\mu_l=0,1,2,3;
\hspace{35pt}
a_{j_l}^{\mu_l}=\left\{ \begin{array}{rl}
             1, & \mu_l=j_l \lor j_l=0\\
             -1, & \mu_l\neq j_l
             \end{array}
   \right.,
\end{equation}
where we used that $\lambda_{0,\ldots,0}$ is always non-negative.
% }}}
\subsection{Another observation from this} % {{{
Choi's matrix of 2 qubits is diagonal in Pauli product basis
(Kraus operators). Does this follow in the general case???? 
% }}}
% }}}
\section{General characterisation of PCE quantum channels} % {{{
\subsubsection*{To avoid confusion} % {{{
\cpnote{Igual el canal totalmente despolarizante y el estado totalmente mixto 
tendrían el mismo diagrama. Quiza el borde se puede ser de diferente color
}
\janote{Me gustaría discutir antes si igual vale la pena hacer la distinción.
Dado que al final la distinción sólo es una ayuda que se me ocurrió para 
explicar gráficamente el procedimiento de la concatenación.}
The figures we've been using (columns, boards, cubes) may represent
both PCE operations (dynamics) and density matrices (kinematics). We
will make a distinction between both representations with colors.
B$\&$W represent PCE operations, and colored figures represent 
density matrices in Pauli products basis.
\begin{figure}
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=3cm]{PCEoperation}
\caption{PCE operation that erases all Pauli components except 
for $r_{0,0}$, $r_{1,1}$, $r_{1,3}$, and $r_{0,3}$.}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=3cm]{state}
\caption{Density matrix in Pauli basis with all components equal
to zero except for $r_{0,0}$, $r_{1,1}$, $r_{1,3}$, and $r_{0,3}$.}
\end{subfigure}
\end{figure}
% }}}

@@

\begin{figure}
  \centering
  \includegraphics[width=.6\textwidth]{cuadro}
\end{figure}
% }}}
% }}}
\section{Things to do and questions to answer} % {{{
\begin{enumerate}
\item This characterization may explain the power-of-2 rule. 
It is quite obvious (in the figures representation) that the concatenation
always erases half of the number of non-zero components. A proof 
for that is missing.
\item Analytic proof for \eqref{eq:inequalities-PCE}. 
\item Analytic proof that \eqref{eq:PCE-characterization} is necessary 
and sufficient to find all elements in $\PCE{n}$.
\item Analytic proof that taking $-1\to 0$ for $a_k^k$ leads to CP.
I suggest using particle swaps and local basis permutation in order 
to reduce the problem to $n$ inequalities.
\item Find a way to count PCE channels.
\end{enumerate}
% }}}

\section{Progress}
\subsection{An idea to prove equation \eqref{eq:eigvals-PCE}}
A density matrix of 1 qubit in Pauli matrices basis is written as
\begin{align}
\rho=\frac{1}{2}\sum_{i=0}^3r_i\sigma_i,
\end{align}
where $\sigma_i$ are a $2\times2$ identity plus Pauli matrices.
For 2 qubits, in the general case, the density matrix cannot be written as
\begin{align}\label{eq:2qubits-densityM-PauliBasis-separable}
\rho&=\frac{1}{4}\qty(\sum_{i=0}^3r_i\sigma_i
\ot
\sum_{j=0}^3r_j\sigma_j)
\end{align}
because a 2-qubits state is not separable, in general. Therefore, 
the density matrix is written as
\begin{align}
\rho=\frac{1}{4}\sum_{i,j=0}^3r_{i,j}\sigma_{i,j}.
\end{align}
In some sense, expanding 
\eqref{eq:2qubits-densityM-PauliBasis-separable}
\begin{align}
\frac{1}{4}\qty(\sum_{i=0}^3r_i\sigma_i
\ot
\sum_{j=0}^3r_j\sigma_j)
&=\frac{1}{4}\qty(
\1_{4}\ot\1_{4}+\1\ot\sum_{j=1}^3r_j\sigma_j+
\sum_{i=1}^3r_i\sigma_i\ot\1+
\sum_{i=1}^3r_i\sigma_i\ot\sum_{j=1}^3r_j\sigma_j
)
\end{align}
makes us a hint realize that, in the general case, the components
of 2-qubits density matrix in tensor products of Pauli matrices basis are
\begin{align}
r_ir_j&\to r_{i.j},&\ r_i&\to r_{i,0},&\ r_j &\to r_{0,j}.
\end{align}
It follows that $n$-quits density matrix is
\begin{align}
\rho=\frac{1}{2^n}\sum_{j_1,\ldots,j_n=0}^3
r_{j_1,\ldots,j_n}\sigma_{j_1}\ot\ldots\ot\sigma_{j_n},
\end{align}
therefore,
\begin{align}
\label{eq:general-comp-of-density-matrix-in-Pauli-matrices-basis}
r_{j_1}\ldots r_{j_k}\ldots r_{j_n}&\to r_{j_1,\ldots,j_n}, &
r_{j_k}&\to r_{0,\ldots,j_k,\ldots,0},
\end{align}
where 
\begin{align}
r_{j_1}\ldots r_{j_k}\ldots r_{j_n}&\neq r_{j_1,\ldots,j_n}, &
r_{j_k}&\neq r_{0,\ldots,j_k,\ldots,0},
\end{align}

Now, the state $\rho_{\E}$ isomorphic to the Choi matrix $D_{\E}$
(is this written ok?) of a 1-qubit PCE channel~$\E$~is
\begin{align}
\rho_{\E}&=\sum_{i=0}^3\lambda_i \dyad{\sigma_i}{\sigma_i},
\end{align}
where $\ket{\sigma_i}$ are the vectorized $\sigma_i$ and
$\lambda_i$ are (this has been shown analytically)
\begin{align}\label{eq:eigvals-1q-choi}
\lambda_0&=\frac{1}{4}\qty(1+\tau_1+\tau_2+\tau_3), &
\lambda_1&=\frac{1}{4}\qty(1+\tau_1-\tau_2-\tau_3),\nonumber\\
\lambda_2&=\frac{1}{4}\qty(1-\tau_1+\tau_2-\tau_3), &
\lambda_3&=\frac{1}{4}\qty(1-\tau_1-\tau_2+\tau_3).
\end{align}

I claim that, in the same sense of 
\eqref{eq:general-comp-of-density-matrix-in-Pauli-matrices-basis},
the eigenvalues of $\rho_{\E}$ of $n$-qubits PCE quantum channel
$\E$ are of the form
\begin{align}
\lambda_{j_1}\ldots \lambda_{j_k}\ldots \lambda_{j_n}
&\to\lambda_{j_1,\ldots,j_n}, &
\lambda_{j_k}&\to\lambda_{0,\ldots,j_k,\ldots,0},
\end{align}
where $\lambda_k$ are the eigenvalues \eqref{eq:eigvals-1q-choi}.

\section{Number of PCE channels by the ratio of components 
left invariant}
\begin{itemize}
\item 1$/$2 of total components invariant:
\begin{align}
4^n-1
\end{align}

\item 1$/$4 of total components invariant:
\begin{align}
\frac{\binom{4^n-1}{2}}{3},
\end{align}

\item 1$/$8 of total components invariant:
\begin{align}
\frac{\binom{4^n-1}{3}-\binom{4^n-1}{2}/3}{28}
\end{align}

\item 1$/$16 of total components invariant:
\begin{align}
\frac{\binom{4^n-1}{4}-35\qty(\frac{\binom{4^n-1}{3}-\binom{4^n-1}{2}/3}{28})}{840}
\end{align}
\end{itemize}

\begin{table}[h!]
\tiny
\begin{tabular}{llllccccccclll}
{\color[HTML]{FF0000} \textbf{1}} &                       &                           &                                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}               & \cellcolor[HTML]{FCE5CD}\textbf{1}   & \textbf{3}                             & \textbf{1}                              & \multicolumn{1}{l}{}                    & \multicolumn{1}{l}{}                &                                                       &                                    &                                \\
{\color[HTML]{FF0000} \textbf{2}} &                       &                           &                                & \multicolumn{1}{l}{} & \cellcolor[HTML]{FCE5CD}\textbf{1} & \cellcolor[HTML]{CFE2F3}\textbf{15}  & \cellcolor[HTML]{FCE5CD}\textbf{35}    & \textbf{15}                             & \textbf{1}                              & \multicolumn{1}{l}{}                &                                                       &                                    &                                \\
{\color[HTML]{FF0000} \textbf{3}} &                       &                           &                                & \textbf{1}           & \textbf{63}                        & \cellcolor[HTML]{FCE5CD}\textbf{651} & \cellcolor[HTML]{CFE2F3}\textbf{1,395} & \cellcolor[HTML]{FCE5CD}\textbf{651}    & \textbf{63}                             & \textbf{1}                          &                                                       &                                    &                                \\
{\color[HTML]{FF0000} \textbf{4}} &                       &                           & \multicolumn{1}{c}{\textbf{1}} & \textbf{255}         & \textbf{10,795}                    & 97,155                               & \cellcolor[HTML]{FCE5CD}200,787        & \cellcolor[HTML]{CFE2F3}\textbf{97,155} & \cellcolor[HTML]{FCE5CD}\textbf{10,795} & \textbf{255}                        & \multicolumn{1}{c}{\textbf{1}}                        &                                    &                                \\
{\color[HTML]{FF0000} \textbf{5}} &                       & \multicolumn{1}{c}{1}     & \multicolumn{1}{c}{1,023}      & 174,251              & 6,347,715                          & 53,743,987                           & \multicolumn{1}{l}{}                   & \cellcolor[HTML]{FCE5CD}53,743,987      & \cellcolor[HTML]{CFE2F3}6,347,715       & \cellcolor[HTML]{FCE5CD}174,251     & \multicolumn{1}{c}{\textbf{1,023}}                    & \multicolumn{1}{c}{\textbf{1}}     &                                \\
{\color[HTML]{FF0000} \textbf{6}} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{4,095} & \multicolumn{1}{c}{2,794,155}  & 408,345,795          & 13,910,980,083                     & \multicolumn{1}{l}{}                 & \multicolumn{1}{l}{}                   & \multicolumn{1}{l}{}                    & \cellcolor[HTML]{FCE5CD}13,910,980,083  & \cellcolor[HTML]{CFE2F3}408,345,795 & \multicolumn{1}{c}{\cellcolor[HTML]{FCE5CD}2,794,155} & \multicolumn{1}{c}{\textbf{4,095}} & \multicolumn{1}{c}{\textbf{1}}
\end{tabular}
\end{table}

\section{2 qubits: hermanitos}
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{hermanitos}
  \caption{Los hermanitos de 2 qubits. La fila de arriba muestra al hermano 
  mayor del canal que tiene exactamente abajo.}
  \label{•}
\end{figure}

>>>>

\section{Eigenvalues of PCE operations' Choi matrix} % {{{
% Intro {{{
Choi matrix's eigenvalues of a 1-qubit Pauli Channel are
\begin{align}
\lambda_{\mu}=1+\sum_{j=1}^3a_j^{\mu}\tau_j
\hspace{35pt}
\mu=0,1,2,3;
\hspace{35pt}
a_j^{\mu}=\left\{ \begin{array}{rl}
             1, & \mu=j \lor \mu=0\\
             -1, & \mu\neq j
             \end{array}
   \right..
\end{align}
Therefore, the inequalities needed to be sattisfied for a 1-qubit 
unital Pauli channel in order to be CP may be expressed in the 
following compact expression:
\begin{equation}
\sum_{j=0}^3a_j^{\mu}\tau_j\geq -1,
\hspace{35pt}
\mu=1,2,3;
\hspace{35pt}
a_j^{\mu}=\left\{ \begin{array}{rl}
             1, & \mu=j\\
             -1, & \mu\neq j
             \end{array}
   \right..
\end{equation}

@@

Therefore, the inequalities needed to be sattisfied for the CP are
\begin{equation}\label{eq:inequalities-PCE}
\sum_{\underset{(j_1,\ldots,j_n\neq0)}{j_1,\ldots,j_n=0}}^3
a_{j_1}^{\mu_1}\ldots a_{j_n}^{\mu_n}\tau_{j_1,\ldots,j_n}
\geq
-1,
\hspace{35pt}
\mu_l=0,1,2,3;
\hspace{35pt}
a_{j_l}^{\mu_l}=\left\{ \begin{array}{rl}
             1, & \mu_l=j_l \lor j_l=0\\
             -1, & \mu_l\neq j_l
             \end{array}
   \right.,
\end{equation}
where we used that $\lambda_{0,\ldots,0}$ is always non-negative.
% }}}
\subsection{Another observation from this} % {{{
Choi's matrix of 2 qubits is diagonal in Pauli product basis
(Kraus operators). Does this follow in the general case???? 
% }}}
% }}}
\section{A general characterization of PCE quantum channels} % {{{
\subsubsection*{To avoid confusion} % {{{
\cpnote{Igual el canal totalmente despolarizante y el estado totalmente mixto 
tendrían el mismo diagrama. Quiza el borde se puede ser de diferente color
}
\janote{Me gustaría discutir antes si igual vale la pena hacer la distinción.
Dado que al final la distinción sólo es una ayuda que se me ocurrió para 
explicar gráficamente el procedimiento de la concatenación.}
The figures we've been using (columns, boards, cubes) may represent
both PCE operations (dynamics) and density matrices (kinematics). We
will make a distinction between both representations with colors.
B$\&$W represent PCE operations, and colored figures represent 
density matrices in Pauli products basis.
\begin{figure}
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=3cm]{PCEoperation}
\caption{PCE operation that erases all Pauli components except 
for $r_{0,0}$, $r_{1,1}$, $r_{1,3}$, and $r_{0,3}$.}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=3cm]{state}
\caption{Density matrix in Pauli basis with all components equal
to zero except for $r_{0,0}$, $r_{1,1}$, $r_{1,3}$, and $r_{0,3}$.}
\end{subfigure}
\end{figure}
% }}}

@@

\begin{figure}
  \centering
  \includegraphics[width=.6\textwidth]{cuadro}
\end{figure}
% }}}
% }}}
\section{Things to do and questions to answer} % {{{
\begin{enumerate}
\item This characterization may explain the power-of-2 rule. 
It is quite obvious (in the figures representation) that the concatenation
always erases half of the number of non-zero components. A proof 
for that is missing. \janote{This is well explained with the vector spaces.}
\item Analytic proof for \eqref{eq:eigvals-PCE}. \janote{Alejo's proof.} 
\item Analytic proof that \eqref{eq:PCE-characterization} is necessary 
and sufficient to find all elements in $\PCE{n}$. \janote{Francois' vector spaces.}
\item Analytic proof that taking $-1\to 0$ for $a_k^k$ leads to CP. \janote{Francois'
general procedure to find PCE channels.}
I suggest using particle swaps and local basis permutation in order 
to reduce the problem to $n$ inequalities.
\item Find a way to count PCE channels. \janote{Also from Francois' vector spaces.}
\end{enumerate}
% }}}

\section{Number of PCE channels by the ratio of components 
left invariant}
%\begin{itemize}
%\item 1$/$2 of total components invariant:
%\begin{align}
%4^n-1
%\end{align}
%
%\item 1$/$4 of total components invariant:
%\begin{align}
%\frac{\binom{4^n-1}{2}}{3},
%\end{align}
%
%\item 1$/$8 of total components invariant:
%\begin{align}
%\frac{\binom{4^n-1}{3}-\binom{4^n-1}{2}/3}{28}
%\end{align}
%
%\item 1$/$16 of total components invariant:
%\begin{align}
%\frac{\binom{4^n-1}{4}-35\qty(\frac{\binom{4^n-1}{3}-\binom{4^n-1}{2}/3}{28})}{840}
%\end{align}
%\end{itemize}
%
%\begin{table}[h!]
%\tiny
%\begin{tabular}{llllccccccclll}
%{\color[HTML]{FF0000} \textbf{1}} &                       &                           &                                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}               & \cellcolor[HTML]{FCE5CD}\textbf{1}   & \textbf{3}                             & \textbf{1}                              & \multicolumn{1}{l}{}                    & \multicolumn{1}{l}{}                &                                                       &                                    &                                \\
%{\color[HTML]{FF0000} \textbf{2}} &                       &                           &                                & \multicolumn{1}{l}{} & \cellcolor[HTML]{FCE5CD}\textbf{1} & \cellcolor[HTML]{CFE2F3}\textbf{15}  & \cellcolor[HTML]{FCE5CD}\textbf{35}    & \textbf{15}                             & \textbf{1}                              & \multicolumn{1}{l}{}                &                                                       &                                    &                                \\
%{\color[HTML]{FF0000} \textbf{3}} &                       &                           &                                & \textbf{1}           & \textbf{63}                        & \cellcolor[HTML]{FCE5CD}\textbf{651} & \cellcolor[HTML]{CFE2F3}\textbf{1,395} & \cellcolor[HTML]{FCE5CD}\textbf{651}    & \textbf{63}                             & \textbf{1}                          &                                                       &                                    &                                \\
%{\color[HTML]{FF0000} \textbf{4}} &                       &                           & \multicolumn{1}{c}{\textbf{1}} & \textbf{255}         & \textbf{10,795}                    & 97,155                               & \cellcolor[HTML]{FCE5CD}200,787        & \cellcolor[HTML]{CFE2F3}\textbf{97,155} & \cellcolor[HTML]{FCE5CD}\textbf{10,795} & \textbf{255}                        & \multicolumn{1}{c}{\textbf{1}}                        &                                    &                                \\
%{\color[HTML]{FF0000} \textbf{5}} &                       & \multicolumn{1}{c}{1}     & \multicolumn{1}{c}{1,023}      & 174,251              & 6,347,715                          & 53,743,987                           & \multicolumn{1}{l}{}                   & \cellcolor[HTML]{FCE5CD}53,743,987      & \cellcolor[HTML]{CFE2F3}6,347,715       & \cellcolor[HTML]{FCE5CD}174,251     & \multicolumn{1}{c}{\textbf{1,023}}                    & \multicolumn{1}{c}{\textbf{1}}     &                                \\
%{\color[HTML]{FF0000} \textbf{6}} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{4,095} & \multicolumn{1}{c}{2,794,155}  & 408,345,795          & 13,910,980,083                     & \multicolumn{1}{l}{}                 & \multicolumn{1}{l}{}                   & \multicolumn{1}{l}{}                    & \cellcolor[HTML]{FCE5CD}13,910,980,083  & \cellcolor[HTML]{CFE2F3}408,345,795 & \multicolumn{1}{c}{\cellcolor[HTML]{FCE5CD}2,794,155} & \multicolumn{1}{c}{\textbf{4,095}} & \multicolumn{1}{c}{\textbf{1}}
%\end{tabular}
%\end{table}

From Francois' manuscript on the PCE's he states that it can be 
proved from the vector spaces' idea that the number of PCE channels 
of $N$ qubits $\mathcal{S}_{N,K}$ that leave $K$ Pauli components invariant is:
\begin{equation}
\mathcal{S}_{N,K}=\prod_{m=0}^{K-1}\frac{2^{2N-m}-1}{2^{K-m}-1}.
\end{equation}

\section{Rainbow correspondence}
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{hermanitos}
  \caption{Los hermanitos de 2 qubits. La fila de arriba muestra al hermano 
  mayor del canal que tiene exactamente abajo.}
  \label{•}
\end{figure}
